Automated Program Verification and Synthesis of Embedded \& Cyber-physical Systems: Advances and Challenges

### 2017-08-17 paper.tex
reviewed section 6
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

Thus, ECPSs demand short development cycles and high level of reliability and robustness~\cite{leeCPS,leeCPS2}, besides presenting challenges that include but are not limited to the ones already imposed by embedded systems. As a consequence of their popularization, human life has also become more and more dependent on services provided by this type of system and, in particular, their success is strictly related to both service relevance and quality/reliability.

@@

\textbf{(RP1)} Cordeiro, Fischer, and Marques-Silva proposed the first SMT-based BMC for full C programs, called Efficient SMT-Based Context-Bounded Model Checker (ESBMC)~\cite{Cordeiro12}, which was later extended to support C++98 programs~\cite{ECBS13}, CUDA programs~\cite{cudalucas,Pereira17}, and Qt-based consumer electronics applications~\cite{Sousa15}. This approach was also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer, in standard benchmarks, which were later confirmed by the benchmarks' creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC)~\cite{CordeiroF11,Cordeiro12}. Other SMT-based BMC approaches have also been proposed and implemented~\cite{MerzFS12}, but the coverage and verification time of all existing ones are still limited to specific program classes, especially for those that contain intensive floating-point arithmetic and dynamic memory allocation~\cite{Beyer14,BeyerSVCOMP15}. One possible research direction is to bridge the gap between BMC tools and SMT solvers, propose background theories, and develop more efficient decision procedures, in order to handle specific program classes. In this particular research direction, the European Project SC$^2$ (Satisfiability Checking and Symbolic Computation: uniting two communities to solve real problems) aims to further extend background theories supported by SMT solvers, in order to solve problems related to security and safety of computer systems (or large mathematical problems), through the development of radically improved software tools~\cite{DBLP:journals/cca/AbrahamA0BBBCDE16}.

\textbf{(RP2)} The SMT-based BMC approach proposed by Cordeiro, Fischer, and Marques-Silva was further developed to verify correct lock acquisition ordering and absence of deadlocks, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira17}, considering monotonic partial-order reduction~\cite{KahlonWG09} and state-hashing techniques, in order to prune state-space exploration~\cite{morse15}. Recent advances for verifying multi-threaded C programs have been proposed to speed up verification times, which significantly prune state-space exploration~\cite{Inverso14,civl15}, and also extensions to verify the Debian GNU/Linux distribution, based on a context- and thread-sensitive abstract interpretation framework~\cite{DBLP:conf/kbse/KroeningPSW16}, were made available; however, the set of concurrent-program classes ({\it e.g.}, CUDA, OpenCL, and MPI) that can be verified is still very limited. One possible research direction is to further extend BMC of multi-threaded C programs via Lazy Sequentialization~\cite{Inverso14}, in order to analyze unsatisfiability cores~\cite{Grumberg05}, with the goal of removing redundant behaviour or analyzing interpolants~\cite{McMillan11} and then prove non-interference of context switches.

@@

{\textbf{(RP5)} Although ESBMC \cite{Cordeiro12} was extended to support C/C++ and some variants, new application interfaces and programming languages are often developed, which require suitable software verification tools. Indeed, it would be interesting if a new programming language model could be loaded, which along with a BMC core could check different programs. Some work towards that was already presented by Sousa, Cordeiro, and Filho~\cite{Garcia16}, which employed operational models for checking Qt-based programs from consumer electronics. In summary, the BMC core (in that case, ESBMC) is not changed (it is still C/C++ code), but instead an operational model, which implements the behavior and features of Qt libraries, is used to provide the new code structure to be checked. This approach could adopt the clang compiler~\cite{Lopes:2014:GSL:2692607}, which is already widely used in industry~\cite{clang-usage}, in order to produce an Abstract Syntax Tree (AST) for a given programming language ({\it e.g.}, C/C++/ObjectiveC/ObjectiveC++). This AST can then be converted either to an intermediate representation (IR) of the underlying verifier or to LLVM bitcode that is already produced by clang. Obviously, a completely new language may present new structures and resources; however, if different inputs are parsed and converted into a form that preserves every aspect of their structure, that can then be analyzed and verified in an unified way. Such research problem is closely related to the first one \textbf{(RP1)} and has the potential to devise a new paradigm in software verification.

{\textbf{(RP6)} State-of-the-art synthesis approaches for embedded (control) systems typically disregard the platform in which the embedded system software operates and restrict themselves to generate code that do not take into account FWL effects. The synthesized system, however, must include the physical plant to avoid serious system's malfunctioning (or even a catastrophe), due to the embedded (control) software, {\it e.g.}, the Mars Polar Lander did not account for leg compressions prior to landing~\cite{Jackson16}. Research in this direction has made some progress to design, implement, and evaluate an automated approach for generating correct-by-construction digital controllers that is based on state-of-the-art inductive synthesis techniques~\cite{Abate17}. Nonetheless, there is still little evidence whether that approach can scale for larger systems modeled by other types of representation ({\it e.g.}, state-space). In addition to that, another research direction for synthesizers is to generate code for UAV trajectory and mission planning by taking into account system's dynamics and nonholonomic constraints~\cite{Araujo16}. As a consequence, verifiers and synthesizers need to handle a wide range of functions, including non-linear and non-convex optimization problems based on both fixed- and floating-point arithmetic.

\textbf{(RP7)} Embedded systems are typically implemented in low- and medium-level programming languages ({\it e.g.}, ANSI-C/C++/Java). Given the availability of different programming languages for writing code for ECPS, the provision of a unified framework able to tackle low- and high-level properties, as well as system behavior, is needed to achieve code coverage in real applications. The clang compiler~\cite{Lopes:2014:GSL:2692607} seems to offer a simple and strong approach to lead software verification to another degree of completeness so that every implementation aspect of an embedded system can be addressed and evaluated. In particular, by using the clang compiler as front-end for a software verification tool, we avoid the need for maintaining a proprietary front-end (a real challenge nowadays, given that the C and C++ standards are rapidly evolving), and then focus on the main objective: formal program verification. Here, we provide only operation models to tackle low- and high-level system properties via an abstract representation of the associated libraries, which conservatively approximates their semantics. The clang compiler offers an industrial quality analyzer and an application programming interface to access and traverse its internal AST, which can then be used by software verifiers to generate their IR or alternatively analyze the LLVM bitcode produced by clang.

>>>>

Thus, ECPSs demand short development cycles and high-level of reliability and robustness~\cite{leeCPS,leeCPS2}, besides presenting challenges that include but are not limited to the ones already imposed by embedded systems. As a consequence of their popularization, human life has also become more and more dependent on services provided by this type of system and, in particular, their success is strictly related to both service relevance and quality/reliability.

@@

\textbf{(RP1)} Cordeiro {\it et al.} proposed the first SMT-based BMC for full C programs, called Efficient SMT-Based Context-Bounded Model Checker (ESBMC)~\cite{Cordeiro12}, which was later extended to support C++98 programs~\cite{ECBS13}, CUDA programs~\cite{cudalucas,Pereira17}, and Qt-based consumer electronics applications~\cite{Sousa15}. This approach was also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer, in standard benchmarks, which were later confirmed by the benchmarks' creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC)~\cite{CordeiroF11,Cordeiro12}. Other SMT-based BMC approaches have also been proposed and implemented~\cite{MerzFS12}, but the coverage and verification time of all existing ones are still limited to specific program classes, especially for those that contain intensive floating-point arithmetic and dynamic memory allocation~\cite{Beyer14,BeyerSVCOMP15}. One possible research direction is to bridge the gap between BMC tools and SMT solvers, propose background theories, and develop more efficient decision procedures, in order to handle specific program classes. In this particular research direction, the European Project SC$^2$ (Satisfiability Checking and Symbolic Computation: uniting two communities to solve real problems) aims to further extend background theories supported by SMT solvers, in order to solve problems related to security and safety of computer systems (or large mathematical problems), through the development of radically improved software tools~\cite{DBLP:journals/cca/AbrahamA0BBBCDE16}.

\textbf{(RP2)} The SMT-based BMC approach proposed by Cordeiro {\it et al.} was further developed to verify correct lock acquisition ordering and absence of deadlocks, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira17}, considering monotonic partial-order reduction~\cite{KahlonWG09} and state-hashing techniques, in order to prune state-space exploration~\cite{morse15}. Recent advances for verifying multi-threaded C programs have been proposed to speed up verification times, which significantly prune state-space exploration~\cite{Inverso14,civl15}, and also extensions to verify the Debian GNU/Linux distribution, based on a context- and thread-sensitive abstract interpretation framework~\cite{DBLP:conf/kbse/KroeningPSW16}, were made available; however, the set of concurrent-program classes ({\it e.g.}, CUDA, OpenCL, and MPI) that can be verified is still very limited. One possible research direction is to further extend BMC of multi-threaded C programs via Lazy Sequentialization~\cite{Inverso14}, in order to analyze unsatisfiability cores~\cite{Grumberg05}, with the goal of removing redundant behaviour or analyzing interpolants~\cite{McMillan11} and then prove non-interference of context switches.

@@

{\textbf{(RP5)} Although ESBMC \cite{Cordeiro12} was extended to support C/C++ and some variants, new application interfaces and programming languages are often developed, which require suitable software verification tools. Indeed, it would be interesting if a new programming language model could be loaded, which along with a BMC core could check different programs. Some work towards that was already presented by Sousa {\it et al.}~\cite{Garcia16}, which employed operational models for checking Qt-based programs from consumer electronics. In summary, the BMC core (in that case, ESBMC) is not changed (it is still C/C++ code), but instead an operational model, which implements the behavior and features of Qt libraries, is used to provide the new code structure to be checked. This approach could adopt the clang compiler~\cite{Lopes:2014:GSL:2692607}, which is already widely used in industry~\cite{clang-usage}, in order to produce an Abstract Syntax Tree (AST) for a given programming language ({\it e.g.}, C/C++/ObjectiveC/ObjectiveC++). This AST can then be converted either to an intermediate representation (IR) of the underlying verifier or to LLVM bitcode that is already produced by clang. Obviously, a completely new language may present new structures and resources; however, if different inputs are parsed and converted into a form that preserves every aspect of their structure, that can then be analyzed and verified in an unified way. Such research problem is closely related to the first one \textbf{(RP1)} and has the potential to devise a new paradigm in ECPS verification.

{\textbf{(RP6)} State-of-the-art synthesis approaches for embedded (control) systems typically disregard the platform in which the embedded system software operates and restrict themselves to generate code that do not take into account FWL effects. The synthesized system, however, must include the physical plant to avoid serious system's malfunctioning (or even a catastrophe), due to the embedded (control) software, {\it e.g.}, the Mars Polar Lander did not account for leg compressions prior to landing~\cite{Jackson16}. Research in this direction has made some progress to design, implement, and evaluate an automated approach for generating correct-by-construction digital controllers that is based on state-of-the-art inductive synthesis techniques~\cite{Abate17,abatecav2017}. Nonetheless, there is still little evidence whether that approach can scale for larger systems modeled by other types of representation ({\it e.g.}, Multiple-Input Multiple-Output). In addition to that, another research direction for synthesizers is to generate code for UAV trajectory and mission planning by taking into account system's dynamics and nonholonomic constraints~\cite{Araujo16}. As a consequence, verifiers and synthesizers need to handle a wide range of functions, including non-linear and non-convex optimization problems based on both fixed- and floating-point arithmetic.

\textbf{(RP7)} Embedded systems are typically implemented in low- and medium-level programming languages ({\it e.g.}, ANSI-C/C++/Java). Given the availability of different programming languages for writing code for ECPS, the provision of a unified framework able to tackle low- and high-level properties, as well as system behavior, is needed to achieve code coverage in real applications. The clang compiler~\cite{Lopes:2014:GSL:2692607} seems to offer a simple and strong approach to lead software verification to another degree of completeness so that every implementation aspect of an ECPS can be addressed and evaluated. In particular, by using the clang compiler as front-end for a software verification tool, we avoid the need for maintaining a proprietary front-end (a real challenge nowadays, given that the C and C++ standards are rapidly evolving), and then focus on the main objective: formal program verification in ECPS. Here, we provide only operation models to tackle low- and high-level system properties via an abstract representation of the associated libraries, which conservatively approximates their semantics. The clang compiler offers an industrial quality analyzer and an application programming interface to access and traverse its internal AST, which can then be used by software verifiers to generate their IR or alternatively analyze the LLVM bitcode produced by clang.
